---
title: "Psychometric Curve Fitting"
output: html_document
---

## Analysis Summary

- Purpose: fit Weibull psychometric functions across lifespan conditions.
- Input metric: ACC by Range from `life_span_updating.csv`.
- Key parameters: Range, ExperimentName (race), Regression, Group, Subject.
- Output plots: Weibull trajectory figures saved to `output/plots/`.
- Output data files: `data/full_data_with_predictions.csv` and `data/full_results.csv`.
- Output console: model convergence status, parameter estimates (lambda, k, JND, PSE, slope), R² values.

# Setup
```{r setup, warning=FALSE}
# Load packages and shared resources
library(tidyverse)
library(minpack.lm) # for nlsLM

# Configure plot text rendering
library(showtext) 
source("R/paths.R")
source("R/theme_pub.R")
font_add("Times New Roman", regular = "times.ttf")
showtext_opts(dpi = 300)
showtext_auto()

# Read and prepare input data
df <- read.csv(data_path("life_span_updating.csv"))


# Remove excluded conditions and bound ACC away from 0/1 for stable fitting
data <- filter(df, Range != 0, !Subject %in% c(229, 241))
data$ACC <- pmin(pmax(data$ACC, 1e-4), 1 - 1e-4)

fit_curve <- function(.data) {
  nlsLM(
    ACC ~ 1 - exp(-(Range / lambda)^k),
    data    = .data,
    start   = list(lambda = 20, k = 1.0),
    lower   = c(lambda = 1e-5, k = 1e-5),
    control = nls.lm.control(
      maxiter = 500  # Increase iterations if needed
    )
  )
}


extract_estimates <- function(model) {

  pars    <- coef(model)
  lambda  <- pars["lambda"]
  k       <- pars["k"]

  j25 <- lambda * (-log(1 - 0.25))^(1 / k)
  j50 <- lambda * (-log(1 - 0.50))^(1 / k)
  j75 <- lambda * (-log(1 - 0.75))^(1 / k)
  JND <- (j75 - j25) / 2
  PSE <- j50
  slope <- (k / (4 * lambda)) * (log(4))^((k - 1) / k)

  # Return Weibull-derived indices in tidy format
  tibble::tibble(
    lambda = lambda,
    k      = k,
    j25    = j25,
    j75    = j75,
    JND    = JND,
    PSE    = PSE,
    slope = slope
  )
}

add_predictions <- function(model, grid) {
  # This function returns predictions for grid according to the model
  grid$Y_hat <- predict(model, newdata = grid)
  grid
}


```
# Fit
```{r data}
# Fit Weibull model per Subject × ExperimentName × Group × Regression
data_fits <- data |> 
  group_by(ExperimentName, Subject,Group,Regression) |> 
  summarise(
    FIT = list(fit_curve(.data = pick(everything())))
  ) |> 
  ungroup()

# Extract Weibull indices from fitted models
data_params <- data_fits |> 
  mutate(
    map(FIT, extract_estimates) |> bind_rows()
  ) |> 
  select(-FIT)
```

Add predictions:

## Prediction Grid and Joined Output

```{r warning=FALSE}
grid <- data.frame(Range = c(15, 21, 27, 33, 39, 45))

# Generate model predictions and join observed data
data_predictions <- data_fits |> 
  group_by(ExperimentName,Subject,Group,Regression) |> 
  reframe(
    add_predictions(FIT[[1]], grid))|> 
  ungroup() |> 
  left_join(data, by = c("ExperimentName", "Subject", "Range","Group","Regression"))%>%
  # Set factor order for plotting
  mutate(
    Group = factor(Group,levels = c("Children","Young Adults", "Middle-Aged", "Elderly")),
    ExperimentName = factor(ExperimentName, levels = c("Caucasian", "Asian"))
  )
```

## Export Model Outputs

Add R2

```{r}
data_params2 <- data_predictions |> 
  group_by(ExperimentName, Subject,Group,Regression) |>
  
  summarise(
    Rsq = cor(ACC, Y_hat) ^ 2
  ) |> 
  ungroup() |> 
  left_join(data_params, by = join_by(ExperimentName, Subject,Group,Regression))

# Save analysis outputs used by downstream scripts
write.csv(data_predictions, data_path("full_data_with_predictions.csv"))
write.csv(data_params2, data_path("full_results.csv"))
```

## Main Weibull Plot

# Weibull plot by race rows and age-group columns (colored by regression)
```{r}

p <- ggplot(data_predictions, aes(x = Range)) +
  facet_grid(rows = vars(ExperimentName), 
             cols = vars(Group),
             labeller = labeller(
               ExperimentName = c("Asian" = "Other-Race", "Caucasian" = "Own-Race")
             )
             ) +
  stat_summary(aes(y = Y_hat, colour = Regression),
               geom = "line", linewidth = 1.2) +
  geom_hline(yintercept = 0.5,linetype = "dashed", linewidth = 0.5, alpha = 0.5) +
  scale_x_continuous(limits = c(15, 45),
                     breaks = c(15, 21, 27, 33, 39, 45)) +
  scale_y_continuous(limits = c(0, 1)) +
  labs(y = "Accuracy",
       x = "Difference in morph levels") +
    theme_pub() +
  scale_colour_manual(
  values = c(biasp = "#7FB3D5", biasm = "#F08080"),
  labels = c(biasp = "Bias+", biasm = "Bias-")
    )

ggsave(filename = plot_path("Weibull_all_groups_lifespan.png"),
       plot  = p,
       width = 10, height = 6, dpi = 300)
```

## Subject-Level Weibull Plots

```{r}
output_dir <- PLOTS_DIR

subjects <- unique(data_predictions$Subject)

for (subj in subjects) {
  
  # Filter to one subject
  dp_subj <- data_predictions %>%
    filter(Subject == subj)
  
  # Build subject-specific plot
  p_subj <- ggplot(dp_subj, aes(x = Range)) +
    facet_grid(
      rows = vars(ExperimentName),
      cols = vars(Group),
      labeller = labeller(
        ExperimentName = c("Asian" = "Other-Race",
                           "Caucasian" = "Own-Race")
      )
    ) +
    stat_summary(aes(y = Y_hat, colour = Regression),
                 geom    = "line",
                 linewidth = 1.2) +
    geom_hline(yintercept = 0.5,
               linetype   = "dashed",
               linewidth  = 0.5,
               alpha      = 0.5) +
    scale_x_continuous(limits = c(15, 45),
                       breaks = c(15, 21, 27, 33, 39, 45)) +
    scale_y_continuous(limits = c(0, 1)) +
    labs(
      y = "Accuracy",
      x = "Difference in morph levels",
      title = paste("Subject", subj)
    ) +
    theme_pub() +
    scale_colour_manual(
      values = c(biasp = "#7FB3D5", biasm = "#F08080"),
      labels = c(biasp = "Bias+", biasm = "Bias-")
    )
  
  # Save subject-specific plot
  ggsave(
    filename = file.path(output_dir,
                         paste0("Subject_", subj, ".png")),
    plot   = p_subj,
    width  = 4, height = 6, dpi = 300
  )
}
```


## Illustration Plots by Group and Race
```{r}
output_dir <- PLOTS_DIR

unique_groups      <- unique(data_predictions$Group)
unique_experiments <- unique(data_predictions$ExperimentName)

for (g in unique_groups) {
  for (e in unique_experiments) {

    subset_df <- data_predictions %>%
    filter(Group == g, ExperimentName == e)
    if (nrow(subset_df) == 0) next

    p <- ggplot(subset_df, aes(x = Range, y = Y_hat, color = Regression)) +
      geom_line(aes(group = interaction(Subject, Regression)), alpha = 0.2) +
      stat_summary(aes(group = Regression), fun = mean, geom = "line", linewidth = 1) +
      geom_hline(yintercept = 0.5, linetype = "dashed", linewidth = 0.5) +
      scale_x_continuous(limits = c(15, 45),
                     breaks = c(15, 21, 27, 33, 39, 45)) +
      scale_y_continuous(limits = c(0, 1)) +
      coord_fixed(ratio = 50) +
      scale_color_manual(
        values = c(biasm = "#F08080", biasp = "#7FB3D5"),
        labels = c(biasp = "Bias+", biasm = "Bias-"),
        name   = "Regression"
      ) +
      labs(
       # title = paste(g, "-", e),
        x     = "Difference in morph levels",
        y     = "Accuracy"
      ) +
      theme_pub()

   filename <- file.path(output_dir, paste0("illustration_option_", g, "_", e, ".png"))
    ggsave(filename,
           plot = p, 
           width = 3,
           height = 4, 
           dpi = 300
           )
    }
  }
```





## Single-Condition Weibull Plots
```{r}
# 1) Get the unique values of your grouping columns
unique_groups       <- unique(data_predictions$Group)
unique_Regressions <- unique(data_predictions$Regression)
unique_experiments <- unique(data_predictions$ExperimentName)

# 2) Loop over all combinations
for (g in unique_groups) {
  for (r in unique_Regressions) {
    for (e in unique_experiments) {
      
      # -- Subset the data for this combination --
      subset_df <- data_predictions %>%
        filter(Group == g, 
               Regression == r, 
               ExperimentName == e)
      
      # If subset_df has no rows, skip
      if(nrow(subset_df) == 0) next
      
      # -- Build the plot --
      p <- ggplot(subset_df, aes(x = Range)) +
        geom_line(aes(y = Y_hat, group = Subject), alpha = 0.15) +
        geom_point(aes(y = ACC), alpha = 0.1, shape = 20) +
        stat_summary(aes(y = Y_hat), geom = "line", color = "red", linewidth = 1) +
        geom_hline(yintercept = 0.5, linetype = "dashed", linewidth = 0.5) +
        coord_cartesian(ylim = c(0, 1)) +
        coord_fixed(ratio = 50) +
        theme_pub() +
        labs(
          title = paste("Group:", g, "- Regression:", r, "- Experiment:", e),
          x = "Range",
          y = "ACC / Y_hat"
        )
      
      # -- Save the plot --
      filename <- plot_path(paste0("plot_", g, "_", r, "_", e, ".png"))
      ggsave(filename = filename, plot = p, width = 8, height = 6)
    }
  }
}

```

## Weibull Plot Colored by Race (Regression Facets) 
```{r}

data_predictions$Regression <- factor(
  data_predictions$Regression,
  levels = c("biasm", "biasp")  # Bias+ on top, Bias− on bottom
)

p <- ggplot(data_predictions, aes(x = Range)) +
  facet_grid(
    rows = vars(Regression), 
    cols = vars(Group),
    labeller = labeller(
      Regression = c("biasm" = "Bias−", "biasp" = "Bias+")
    )
  ) +
  stat_summary(
    aes(y = Y_hat, colour = ExperimentName),
    geom = "line", linewidth = 1.2
  ) +
  geom_hline(yintercept = 0.5, linetype = "dashed", linewidth = 0.5, alpha = 0.5) +
  scale_x_continuous(
    limits = c(15, 45),
    breaks = c(15, 21, 27, 33, 39, 45)
  ) +
  scale_y_continuous(limits = c(0, 1)) +
  labs(
    y = "Accuracy",
    x = "Difference in morph levels"
  ) +
  theme_pub() +
  scale_colour_manual(
    values = c("Caucasian" = "#A6D8C3", "Asian" = "#F6B48F"),
    labels = c(Caucasian = "Own-Race", Asian = "Other-Race")
  )

ggsave(
  filename = plot_path("Weibull_all_groups_lifespan_switched.png"),
  plot = p,
  width = 10, height = 6, dpi = 300
)



```



